Points Above Replacement: A New NBA Metric to Evaluate Player Performance

### 2022-04-21 Manuscript/Manuscript.tex
Add files via upload

\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
\captionsetup{labelformat=simple, labelsep=period}
%\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
%\renewcommand{\theequation}{\arabic{equation}}
%\makeatletter
%\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}

@@

   \end{center}
\end{titlepage}
% Abstracts are required.
\section*{Abstract}
In attempting to predict the score of an NBA game, players who are 
injured, suspended, or for any other 
reason not playing have a significant impact on said prediction. Baseball 
uses a statistic called Wins Above 
Replacement, or WAR, to tell how many wins a team would gain or lose if 
they replaced a certain player 
with one who is league average. The goal of this paper is to create a 
similar statistic, called Points Above 
Replacement, or PAR, that uses statistics from throughout the season to 
predict how many points a team 
gains or loses by missing a certain player, and thus having to replace him 
with one who is "league 
average". The statistic is standardized and adjusted by both position and 
value to a player's given team. 
The model roughly follows the calculation of WAR in baseball, assigning 
values to points created on offense 
and defense, as well as a positional and team adjustment. A comparison 
with the created metric along with 
other well-known advanced player metrics is done via a multiple 
regression.

@@

The basic method for calculating the PAR method was based on how 
WAR is calculated in baseball, but 
also took influences from other advanced basketball statistics. WAR is 
based on Offensive and Defensive 
Runs, which are calculated subjectively via attributes of the game such as 
hitting, base running, fielding, 
and so forth. The subjectivity is in the coefficients chosen to scale 
statistics which impact Offensive and 
Defensive Runs. Thus, there are actually an infinite different kinds of WAR 
in baseball, depending on which 
method of calculating Offensive and Defensive Runs is used. Similarly, 
PAR is created by calculating 
Offensive and Defensive Points statistics, to analyze how many points a 
player adds for his team, and how 
many he prevents the other team from scoring. Using the basic offensive 
statistics of Effective Field Goal 
Percentage, Three Point Percentage, Free Throw Percentage, Offensive 
Rebounds, Assists, Points, and 
Turnovers, we assign a coefficient to them based on how important they 
are for a given position. Point 
guards are expected to pass more than centers, and a point guard's 
passing ability has more value to his 
team than a center, so more weight is placed on that statistic for point 
guards. This goes for all positions, 
and a coefficient matrix of a constant scale is created. That is to say, 
assists have a coefficient from point 
guards, shooting guards, small forwards, power forwards, and centers, 
respectively, of 4, 3.5, 3, 2.5, and 2. 
The coefficient by position method is one similar to that of Box Plus 
Minus, another advanced NBA metric. 
The position a player was assigned for the calculation was the position 
where he played the most minutes 
throughout the 2020-2021 season. This coefficient method gives a clean 
scaling into the calculation of 
Offensive Points. Some statistics, such as Points and Effective Field Goal 
Percentage, are equally 
important, regardless of the position of a given player. Thus, these 
statistics have a constant coefficient that 
does not change. Once all coefficients were applied to all statistics, the 
sum was taken to give Offensive 
Points. The exact same method was applied to the Defensive Points 
calculation, but used the basic 
defensive statistics of Defensive Rebounds, Total Rebounds, Steals, 
Blocks, and Personal Fouls. As 
turnovers and personal fouls are not desirable amongst players, the 
coefficients for these statistics is 
negative, which allows for a scale of not only the good a player does, but 
also what they do poorly. This 
coefficient method is similar to the Positional Adjustment metric applied in 
WAR calculation, which seeks to 
adjust a player's importance based on what is expected of their position. 
The coefficients for each position are listed in Table~
\ref{tab:Coefficientstable}. Coefficients are subjective, but were changed 
to have a constant scale by position, and create a PAR scale that lined up 
closely with other advanced metrics.

\begin{table}[H]
  \caption{Coefficients for Each Statistic in Offensive and Defensive Points}
  \label{tab:Coefficientstable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Statistic & Point & Shooting & Small & Power  & Center\\
  & Guard & Guard & Forward & Forward & \\
  \midrule
 Effective FG Percentage & 200 & 200 & 200 & 200 & 200\\
 3 Point Percentage & 300 & 275 & 250 & 225 & 200\\
 Free Throw Percentage & 100 & 90 & 80 & 70 & 60\\
 Offensive Rebounds & 1 & 1.5 & 2 & 2.5 & 3\\
 Assists & 4 & 3.5 & 3 & 2.5 & 2\\
 Points & 2 & 2 & 2 & 2 & 2\\
 Turnovers & -5 & -4.5 & -4 & -3.5 & -3\\
 Defensive Rebounds & 2 & 2.5 & 3 & 3.5 & 4\\
 Total Rebounds & 3 & 3.25 & 3.5 & 3.75 & 4\\
 Steals & 5 & 4.5 & 4 & 3.5 & 3\\
 Blocks & 3 & 3.5 & 4 & 4.5 & 5\\
 Personal Fouls & -3 & -3 & -3 & -3 & -3\\
  \bottomrule
\end{tabular}
\end{table}

The next step was creating a Team Adjustment statistic. As stated before, 
the PAR metric seeks to 
anticipate how valuable a player is to their given team, rather than as an 
average of the entire league. This 
gives more of an insight of how a team will fare without a certain player. 
We expect this statistic to change if 
a player were to switch teams. This is similar to the League Adjustment in 
WAR calculation, where a 
player's value is adjusted based on whether he plays in the American 
League or the National League. The 
Eastern and Western Conferences in the NBA have much less disparity 
compared to the MLB, and each 
conference plays each other much more frequently, so we instead change 
this to adjust a player's value to 
his specific team. The team assigned to a given player was the team 
where he played the most games 
during the 2020-2021 season. If a player had played an even amount of 
games for 2 separate teams, the 
team the player had played the most minutes for was assigned as his 
team. The Team Adjustment statistic 
is then calculated by taking the total minutes a player was on the court 
during the entire season, and diving 
it by a constant of 3,456 (the maximum number of minutes a player could 
have played during the shortened 
season, 48 minutes per game multiplied by 72 games). Then, the 
percentage of points a player scored for 
his team (player's total points divided by team's total points throughout the 
season) was calculated, and the 
two results were multiplied together. As a player cannot play 100 percent 
of his team's minutes, or score 
100 percent of his team's points, the result is a number between 0 and 1. 
A similar adjustment is also used 
in the calculation of Box Plus Minus, which is a key factor in calculating 
Value Over Replacement Player, 
two of the advanced metrics PAR will be compared against. The goal of 
this adjustment is to give less of a 
penalty to players who play a majority of their team's minutes, and score a 
majority of their team's points. 
This helps truly elite players stand out amongst the pack, as well as 
players who may not be the best of the 
best, but matter more to their team specifically than in the grand scheme 
of things.

@@

\begin{table}[H]
  \caption{PAR Values from the Entire NBA}
  \label{tab:NBAtable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Position & Team & PAR\\
  \midrule
 1 & Nikola Jokic & C & Denver Nuggets & 7.69\\
 2 & Nikola Vucevic & C & Orlando Magic & 6.55\\
 3 & Julius Randle & PF & New York Knicks & 6.49\\
 4 & Luka Doncic & PG & Dallas Mavericks & 4.61\\
 5 & Damian Lillard & PG & Portland Trail Blazers & 4.47\\
 6 & Giannis Antetokounmpo & PF & Milwaukee Bucks & 4.21\\
 7 & Stephen Curry & PG & Golden State Warriors & 4.17\\
 8 & Russell Westbrook & PG & Washington Wizards & 4.05\\
 9 & Jayson Tatum & SF & Boston Celtics & 3.96\\
 10 & Domanatas Sabonis & PF & Indiana Pacers & 3.21\\
 \bottomrule
\end{tabular}
\end{table}

As can be seen in Table~\ref{tab:NBAtable}, one player rises above all others in 
both PAR and VORP. 
That player is Nikola Jokic of the 
Denver Nuggets, and this "outlier" makes perfect sense, and helps show 
just how dominant his MVP 
season was. Not only did Jokic average 26.4 points, 10.8 rebounds, and 
8.3 assists per game, he was 
doing it from the center position. Having a center that has the ability to 
perform in all aspects of the game, 
even those that his position does not require, proves his value to his 
team, which is why his value is so 
much higher in both the new PAR metric, and the accepted VORP metric. 
This logic applies to Nikola 
Vucevic, Julius Randle, Russell Westbrook, and Domanatas Sabonis as 
well. While they might not be 
considered the best at their position, or an extremely elite player in the 
NBA, they are very versatile players, 
that can do anything on the court, and thus they are rewarded with a high 
PAR value. This makes sense, as 
without these players, a lot of the statistics their team accumulates are 
gone, so while they may not have as 
much value in the grand scheme of the NBA, they were very much the 
most valuable players for their 
teams. Another key thing to note is that some players, such as LeBron 
James, have a lower PAR than 
expected, and this could be due to the fact that they played most 
frequently in a position that did not match 
their skill set. LeBron James played primarily as a point guard, but his 
skillset matches that of a small 
forward or power forward. Thus, his strengths are not adequately 
measured, as he is essentially being 
played "out of position".

\begin{table}[H]
  \caption{PAR Values from all Point Guards}
  \label{tab:PGtable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Luka Doncic & Dallas Mavericks & 4.61\\
 2 & Damian Lillard & Portland Trail Blazers & 4.47\\
 3 & Stephen Curry & Golden State Warriors & 4.17\\
 4 & Russell Westbrook & Washington Wizards & 4.05\\
 5 & Trae Young & Atlanta Hawks & 2.69\\
 6 & De'Aaron Fox & Sacramento Kings & 1.94\\
 7 & Chris Paul & Phoenix Suns & 1.93\\
 8 & Kyrie Irving & Brooklyn Nets & 1.74\\
 9 & Dejounte Murray & San Antonio Spurs & 1.64\\
 10 & Ja Morant & Memphis Grizzlies & 1.45\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Shooting Guards}
  \label{tab:SGtable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Bradley Beal & Washington Wizards & 3.17\\
 2 & Devin Booker & Phoenix Suns & 2.74\\
 3 & Terry Rozier & Charlotte Hornets & 2.59\\
 4 & Zach Lavine & Chicago Bulls & 2.29\\
 5 & RJ Barrett & New York Knicks & 2.28\\
 6 & Anthony Edwards & Minnesota Timberwolves & 2.22\\
 7 & Jaylen Brown & Boston Celtics & 1.87\\
 8 & Collin Sexton & Cleveland Cavaliers & 1.85\\
 9 & Buddy Hield & Sacramento Kings & 1.72\\
 10 & Norman Powell & Toronto Raptors & 1.37\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Small Forwards}
  \label{tab:SFtable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Jayson Tatum & Boston Celtics & 3.96\\
 2 & Khris Middleton & Milwaukee Bucks & 2.39\\
 3 & Brandon Ingram & New Orleans Pelicans & 2.13\\
 4 & Kawhi Leonard & Los Angeles Clippers & 1.58\\
 5 & Paul George & Los Angeles Clippers & 1.44\\
 6 & Michael Porter Jr. & Denver Nuggets & 1.42\\
 7 & Jimmy Butler & Miami Heat & 1.41\\
 8 & Bojan Bogdonavic & Utah Jazz & 1.35\\
 9 & Mikal Bridges & Phoenix Suns & 1.19\\
 10 & Jerami Grant & Detroit Pistons & 1.14\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Power Forwards}
  \label{tab:PFtable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Julius Randle & New York Knicks & 6.49\\
 2 & Giannis Antetokounmpo & Milwaukee Bucks & 4.21\\
 3 & Domanatas Sabonis & Indiana Pacers & 3.21\\
 4 & Zion Williamson & New Orleans Pelicans & 3.07\\
 5 & Andrew Wiggins & Golden State Warriors & 2.26\\
 6 & Demar Derozan & San Antonio Spurs & 1.88\\
 7 & Tobias Harris & Philadelphia 76ers & 1.87\\
 8 & Pascal Siakam & Toronto Raptors & 1.72\\
 9 & John Collins & Atlanta Hawks & 1.35\\
 10 & Harrison Barnes & Sacramento Kings & 1.11\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Centers}
  \label{tab:Ctable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Nikola Jokic & Denver Nuggets & 7.69\\
 2 & Nikola Vucevic & Orlando Magic & 6.55\\
 3 & Rudy Gobert & Utah Jazz & 3.10\\
 4 & Bam Adebayo & Miami Heat & 2.78\\
 5 & Joel Embiid & Philadelphia 76ers & 2.26\\
 6 & Clint Capela & Atlanta Hawks & 2.25\\
 7 & DeAndre Ayton & Phoenix Suns & 2.11\\
 8 & Jonas Valanciunas & Memphis Grizzlies & 1.98\\
 9 & Karl-Anthony Towns & Minnesota Timberwolves & 1.82\\
 10 & Kelly Olynyk & Miami Heat & 1.39\\
  \bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[H]
  \caption{Multiple Regression on PAR}
  \label{tab:MultipleRegression}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Variable &  Estimate & Standard Error & P-Value\\
  \midrule
Intercept & -1.009821  & 0.144907 & 9.44 * e-12\\
PER & 0.030409 & 0.009569 & 0.00157\\
WS & 0.121319 & 0.023793 & 4.75 * e-7\\
BPM & -0.056802 & 0.012889 & 1.27 * e-5\\
VORP & 0.558947 & 0.048688 & Less Than 2 * e-16\\
  \bottomrule
\end{tabular}
\end{table}

Not included in the table Table~\ref{tab:MultipleRegression} is the adjusted 
R-squared 
value of 0.6974. This 
statistic is key, as it shows that 69.74 percent of variation 
in PAR can be explained by these other 4 accepted metrics. This shows a 
clear correlation between PAR 
and the metrics, but also shows that PAR has something unique to offer.

\section{Discussion}
% Analyzing entire study, show how you can use PAR now, did it achieve the overall goal I was looking for?
% Talk about drawbacks, what could be improved, etc.
% Talk about how it is in agreement with existing measures
Ultimately, the goal of this statistic is to allow a reader to gauge a player's 
value to his team based on a 
league average player. As previously stated, the metric is adjusted so that 
a league average player has a 
PAR value of 0. Based on all other players' PAR values, we can create a 
sort of "legend", a rough guideline 
for what a player's role will be based on the threshold his PAR value falls 
under.

\begin{table}[H]
  \caption{PAR Legend}
  \label{tab:PARLegend}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Role &  PAR\\
  \midrule
Benchwarmer & Less than -0.43\\
Role Player & -0.43 to 0.08\\
Starter & 0.08 to 1.09\\
All-Star & 1.09 to 1.94\\
Superstar & Greater than 1.94\\
  \bottomrule
\end{tabular}
\end{table}

This legend in Table~\ref{tab:PARLegend} was calculated based on a quantile 
calculation of all PAR 
data. It is not an even distribution, as 
it is assumed that there are many more benchwarmers in the NBA than 
there are superstars. Thus, 
benchwarmers are defined as the bottom 50 percent of the league, as an 
NBA roster of about 15 players 
will only give 7 or 8 players significant minutes. Players between the 50th 
and 75th percentile are defined 
as role players, as they get significant minutes, but do not make an impact 
enough to be considered a solid 
starter. This is where  a "league average" player is expected to fall, as 0 is 
contained in this part of the 
legend, and 0 represents a league average player via this metric. Next, 
the 75th to 90th percentile are 
defined as starters, because only 5 players can start a game, and, more 
often than not, a team only has 3 
or 4 players who are guaranteed starters. 90th to 95th percentiles 
represent all-stars, as that is clearly a 
very prestigious area to be a part of. Lastly, it is assumed that only the top 
5 percent of players in the 
league can be deemed "superstars", so they comprise the 95th to 100th 
percentile of the NBA. By this 
method, it is felt that a more realistic legend for PAR has been created as 
opposed to just breaking up the 
data evenly into 5 groups.

>>>>

%\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
%\captionsetup{labelformat=simple, labelsep=period}
%\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
%\renewcommand{\theequation}{\arabic{equation}}
%\makeatletter
%\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}

@@

   \end{center}
\end{titlepage}
% Abstracts are required.
\section*{Abstract}
In attempting to predict the score of an NBA game, players who are 
injured, suspended, or for any other 
reason not playing have a significant impact on said prediction. Baseball 
uses a statistic called Wins Above 
Replacement, or WAR, to tell how many wins a team would gain or lose if 
they replaced a certain player 
with one who is league average. The goal of this paper is to create a 
similar statistic, called Points Above 
Replacement, or PAR, that uses statistics from throughout the season to 
predict how many points a team 
gains or loses by missing a certain player, and thus having to replace him 
with one who is ``league 
average''. The statistic is standardized and adjusted by both position and 
value to a player's given team. 
The model roughly follows the calculation of WAR in baseball, assigning 
values to points created on offense 
and defense, as well as a positional and team adjustment. A comparison 
with the created metric along with 
other well-known advanced player metrics is done via a multiple 
regression.

@@

The basic method for calculating the PAR method was based on how 
WAR is calculated in baseball, but 
also took influences from other advanced basketball statistics. WAR is 
based on Offensive and Defensive 
Runs, which are calculated subjectively via attributes of the game such as 
hitting, base running, fielding, 
and so forth \citep{BaseballWAR}. The subjectivity is in the coefficients 
chosen to scale 
statistics which impact Offensive and 
Defensive Runs. Thus, there are actually an infinite different kinds of WAR 
in baseball, depending on which 
method of calculating Offensive and Defensive Runs is used. Similarly, 
PAR is created by calculating 
Offensive and Defensive Points statistics, to analyze how many points a 
player adds for his team, and how 
many he prevents the other team from scoring.

Using the basic offensive 
statistics of Effective Field Goal 
Percentage, Three Point Percentage, Free Throw Percentage, Offensive 
Rebounds, Assists, Points, and 
Turnovers, we assign a coefficient to them based on how important they 
are for a given position. Point 
guards are expected to pass more than centers, and a point guard's 
passing ability has more value to his 
team than a center, so more weight is placed on that statistic for point 
guards. This goes for all positions, 
and a coefficient matrix of a constant scale is created. That is to say, 
assists have a coefficient from point 
guards, shooting guards, small forwards, power forwards, and centers, 
respectively, of 4, 3.5, 3, 2.5, and 2.

The coefficient by position method is one similar to that of Box Plus 
Minus, another advanced NBA metric \citep{BasketballReferenceBPM}. 
The position a player was assigned for the calculation was the position 
where he played the most minutes 
throughout the 2020-2021 season. This coefficient method gives a clean 
scaling into the calculation of 
Offensive Points. Some statistics, such as Points and Effective Field Goal 
Percentage, are equally 
important, regardless of the position of a given player. Thus, these 
statistics have a constant coefficient that 
does not change. Once all coefficients were applied to all statistics, the 
sum was taken to give Offensive 
Points.

The exact same method was applied to the Defensive Points 
calculation, but used the basic 
defensive statistics of Defensive Rebounds, Total Rebounds, Steals, 
Blocks, and Personal Fouls. As 
turnovers and personal fouls are not desirable amongst players, the 
coefficients for these statistics is 
negative, which allows for a scale of not only the good a player does, but 
also what they do poorly. This 
coefficient method is similar to the Positional Adjustment metric applied in 
WAR calculation, which seeks to 
adjust a player's importance based on what is expected of their position.

The coefficients for each position are listed in Table~
\ref{tab:Coefficientstable}. Coefficients are subjective, but were changed 
to have a constant scale by position, and create a PAR scale that lined up 
closely with other advanced metrics. This coefficient matrix could be 
looked at as an optimization problem, where coefficients by position are 
changed to achieve as high of a correlation with other advanced metrics 
as possible. However, a constant scale by position and a relatively high 
correlation to other metrics shows uniqueness in the statistic but also 
validity.

\begin{table}[H]
  \caption{Coefficients for Each Statistic in Offensive and Defensive Points}
  \label{tab:Coefficientstable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Statistic & Point & Shooting & Small & Power  & Center\\
  & Guard & Guard & Forward & Forward & \\
  \midrule
 Effective FG Percentage & 200 & 200 & 200 & 200 & 200\\
 3 Point Percentage & 300 & 275 & 250 & 225 & 200\\
 Free Throw Percentage & 100 & 90 & 80 & 70 & 60\\
 Offensive Rebounds & 1 & 1.5 & 2 & 2.5 & 3\\
 Assists & 4 & 3.5 & 3 & 2.5 & 2\\
 Points & 2 & 2 & 2 & 2 & 2\\
 Turnovers & $-5$ & $-4.5$ & $-4$ & $-3.5$ & $-3$\\
 Defensive Rebounds & 2 & 2.5 & 3 & 3.5 & 4\\
 Total Rebounds & 3 & 3.25 & 3.5 & 3.75 & 4\\
 Steals & 5 & 4.5 & 4 & 3.5 & 3\\
 Blocks & 3 & 3.5 & 4 & 4.5 & 5\\
 Personal Fouls & $-3$ & $-3$ & $-3$ & $-3$ & $-3$\\
  \bottomrule
\end{tabular}
\end{table}

The next step was creating a Team Adjustment statistic. As stated before, 
the PAR metric seeks to 
anticipate how valuable a player is to their given team, rather than as an 
average of the entire league. This 
gives more of an insight of how a team will fare without a certain player. 
We expect this statistic to change if 
a player were to switch teams. This is similar to the League Adjustment in 
WAR calculation, where a 
player's value is adjusted based on whether he plays in the American 
League or the National League. The 
Eastern and Western Conferences in the NBA have much less disparity 
compared to the MLB, and each 
conference plays each other much more frequently, so we instead change 
this to adjust a player's value to 
his specific team. The team assigned to a given player was the team 
where he played the most games 
during the 2020-2021 season. If a player had played an even amount of 
games for 2 separate teams, the 
team the player had played the most minutes for was assigned as his 
team.

The Team Adjustment statistic 
is then calculated by taking the total minutes a player was on the court 
during the entire season, and diving 
it by a constant of 3,456 (the maximum number of minutes a player could 
have played during the shortened 
season, 48 minutes per game multiplied by 72 games). Then, the 
percentage of points a player scored for 
his team (player's total points divided by team's total points throughout the 
season) was calculated, and the 
two results were multiplied together. As a player cannot play 100 percent 
of his team's minutes, or score 
100 percent of his team's points, the result is a number between 0 and 1.

A similar adjustment is also used 
in the calculation of Box Plus Minus, which is a key factor in calculating 
Value Over Replacement Player, 
two of the advanced metrics PAR will be compared against. The goal of 
this adjustment is to give less of a 
penalty to players who play a majority of their team's minutes, and score a 
majority of their team's points. 
This helps truly elite players stand out amongst the pack, as well as 
players who may not be the best of the 
best, but matter more to their team specifically than in the grand scheme 
of things.

@@

\begin{table}[H]
  \caption{PAR Values from the Entire NBA}
  \label{tab:NBAtable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Position & Team & PAR\\
  \midrule
 1 & Nikola Jokic & C & Denver Nuggets & 7.69\\
 2 & Nikola Vucevic & C & Orlando Magic & 6.55\\
 3 & Julius Randle & PF & New York Knicks & 6.49\\
 4 & Luka Doncic & PG & Dallas Mavericks & 4.61\\
 5 & Damian Lillard & PG & Portland Trail Blazers & 4.47\\
 6 & Giannis Antetokounmpo & PF & Milwaukee Bucks & 4.21\\
 7 & Stephen Curry & PG & Golden State Warriors & 4.17\\
 8 & Russell Westbrook & PG & Washington Wizards & 4.05\\
 9 & Jayson Tatum & SF & Boston Celtics & 3.96\\
 10 & Domanatas Sabonis & PF & Indiana Pacers & 3.21\\
 \bottomrule
\end{tabular}
\end{table}

As can be seen in Table~\ref{tab:NBAtable}, one player rises above all 
others in 
both PAR and VORP. 
That player is Nikola Jokic of the 
Denver Nuggets, and this ``outlier'' makes perfect sense, and helps show 
just how dominant his MVP 
season was. Not only did Jokic average 26.4 points, 10.8 rebounds, and 
8.3 assists per game, he was 
doing it from the center position. Having a center that has the ability to 
perform in all aspects of the game, 
even those that his position does not require, proves his value to his 
team, which is why his value is so 
much higher in both the new PAR metric, and the accepted VORP metric. 
This logic applies to Nikola 
Vucevic, Julius Randle, Russell Westbrook, and Domanatas Sabonis as 
well. While they might not be 
considered the best at their position, or an extremely elite player in the 
NBA, they are very versatile players, 
that can do anything on the court, and thus they are rewarded with a high 
PAR value. This makes sense, as 
without these players, a lot of the statistics their team accumulates are 
gone, so while they may not have as 
much value in the grand scheme of the NBA, they were very much the 
most valuable players for their 
teams. Another key thing to note is that some players, such as LeBron 
James, have a lower PAR than 
expected, and this could be due to the fact that they played most 
frequently in a position that did not match 
their skill set. LeBron James played primarily as a point guard, but his 
skillset matches that of a small 
forward or power forward. Thus, his strengths are not adequately 
measured, as he is essentially being 
played ``out of position''.

\begin{table}[H]
  \caption{PAR Values from all Point Guards}
  \label{tab:PGtable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Luka Doncic & Dallas Mavericks & 4.61\\
 2 & Damian Lillard & Portland Trail Blazers & 4.47\\
 3 & Stephen Curry & Golden State Warriors & 4.17\\
 4 & Russell Westbrook & Washington Wizards & 4.05\\
 5 & Trae Young & Atlanta Hawks & 2.69\\
 6 & De'Aaron Fox & Sacramento Kings & 1.94\\
 7 & Chris Paul & Phoenix Suns & 1.93\\
 8 & Kyrie Irving & Brooklyn Nets & 1.74\\
 9 & Dejounte Murray & San Antonio Spurs & 1.64\\
 10 & Ja Morant & Memphis Grizzlies & 1.45\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Shooting Guards}
  \label{tab:SGtable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Bradley Beal & Washington Wizards & 3.17\\
 2 & Devin Booker & Phoenix Suns & 2.74\\
 3 & Terry Rozier & Charlotte Hornets & 2.59\\
 4 & Zach Lavine & Chicago Bulls & 2.29\\
 5 & RJ Barrett & New York Knicks & 2.28\\
 6 & Anthony Edwards & Minnesota Timberwolves & 2.22\\
 7 & Jaylen Brown & Boston Celtics & 1.87\\
 8 & Collin Sexton & Cleveland Cavaliers & 1.85\\
 9 & Buddy Hield & Sacramento Kings & 1.72\\
 10 & Norman Powell & Toronto Raptors & 1.37\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Small Forwards}
  \label{tab:SFtable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Jayson Tatum & Boston Celtics & 3.96\\
 2 & Khris Middleton & Milwaukee Bucks & 2.39\\
 3 & Brandon Ingram & New Orleans Pelicans & 2.13\\
 4 & Kawhi Leonard & Los Angeles Clippers & 1.58\\
 5 & Paul George & Los Angeles Clippers & 1.44\\
 6 & Michael Porter Jr. & Denver Nuggets & 1.42\\
 7 & Jimmy Butler & Miami Heat & 1.41\\
 8 & Bojan Bogdonavic & Utah Jazz & 1.35\\
 9 & Mikal Bridges & Phoenix Suns & 1.19\\
 10 & Jerami Grant & Detroit Pistons & 1.14\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Power Forwards}
  \label{tab:PFtable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Julius Randle & New York Knicks & 6.49\\
 2 & Giannis Antetokounmpo & Milwaukee Bucks & 4.21\\
 3 & Domanatas Sabonis & Indiana Pacers & 3.21\\
 4 & Zion Williamson & New Orleans Pelicans & 3.07\\
 5 & Andrew Wiggins & Golden State Warriors & 2.26\\
 6 & Demar Derozan & San Antonio Spurs & 1.88\\
 7 & Tobias Harris & Philadelphia 76ers & 1.87\\
 8 & Pascal Siakam & Toronto Raptors & 1.72\\
 9 & John Collins & Atlanta Hawks & 1.35\\
 10 & Harrison Barnes & Sacramento Kings & 1.11\\
  \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
  \caption{PAR Values from all Centers}
  \label{tab:Ctable}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Rank & Player & Team & PAR\\
  \midrule
 1 & Nikola Jokic & Denver Nuggets & 7.69\\
 2 & Nikola Vucevic & Orlando Magic & 6.55\\
 3 & Rudy Gobert & Utah Jazz & 3.10\\
 4 & Bam Adebayo & Miami Heat & 2.78\\
 5 & Joel Embiid & Philadelphia 76ers & 2.26\\
 6 & Clint Capela & Atlanta Hawks & 2.25\\
 7 & DeAndre Ayton & Phoenix Suns & 2.11\\
 8 & Jonas Valanciunas & Memphis Grizzlies & 1.98\\
 9 & Karl-Anthony Towns & Minnesota Timberwolves & 1.82\\
 10 & Kelly Olynyk & Miami Heat & 1.39\\
  \bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[H]
  \caption{Multiple Regression on PAR}
  \label{tab:MultipleRegression}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Variable &  Estimate & Standard Error & P-Value\\
  \midrule
Intercept & $-1.009$  & 0.145 & $9.44 \times 10^{-12}$\\
PER & 0.030 & 0.009 & 0.002\\
WS & 0.121 & 0.024 & $4.75 \times 10^{-7}$\\
BPM & $-0.057$ & 0.013 & $1.27 \times 10^{-5}$\\
VORP & 0.559 & 0.049 & Less Than $2 \times 10^{-16}$\\
  \bottomrule
\end{tabular}
\end{table}

In Table~\ref{tab:MultipleRegression} an adjusted R-squared value of 
0.6974 is shown, along with an SER of 0.5501. The adjusted R-squared
statistic is key, as it shows that 69.74 percent of variation 
in PAR can be explained by these other 4 accepted metrics. This shows a 
clear correlation between PAR 
and the metrics, but also shows that PAR has something unique to offer.

\section{Discussion}
% Analyzing entire study, show how you can use PAR now, did it achieve the overall goal I was looking for?
% Talk about drawbacks, what could be improved, etc.
% Talk about how it is in agreement with existing measures
Ultimately, the goal of this statistic is to allow a reader to gauge a player's 
value to his team based on a 
league average player. As previously stated, the metric is adjusted so that 
a league average player has a 
PAR value of 0. Based on all other players' PAR values, we can create a 
sort of ``legend'', a rough guideline 
for what a player's role will be based on the threshold his PAR value falls 
under.

\begin{table}[H]
  \caption{PAR Legend}
  \label{tab:PARLegend}
\centering
\begin{tabular}[t]{lcccccc}
  \toprule
  Role &  PAR\\
  \midrule
Benchwarmer & Less than $-0.43$\\
Role Player & $-0.43$ to 0.08\\
Starter & 0.08 to 1.09\\
All-Star & 1.09 to 1.94\\
Superstar & Greater than 1.94\\
  \bottomrule
\end{tabular}
\end{table}

This legend in Table~\ref{tab:PARLegend} was calculated based on a 
quantile 
calculation of all PAR 
data. It is not an even distribution, as 
it is assumed that there are many more benchwarmers in the NBA than 
there are superstars. Thus, 
benchwarmers are defined as the bottom 50 percent of the league, as an 
NBA roster of about 15 players 
will only give 7 or 8 players significant minutes. Players between the 50th 
and 75th percentile are defined 
as role players, as they get significant minutes, but do not make an impact 
enough to be considered a solid 
starter. This is where  a `league average'' player is expected to fall, as 0 is 
contained in this part of the 
legend, and 0 represents a league average player via this metric. Next, 
the 75th to 90th percentile are 
defined as starters, because only 5 players can start a game, and, more 
often than not, a team only has 3 
or 4 players who are guaranteed starters. The 90th to 95th percentiles 
represent all-stars, as that is clearly a 
very prestigious area to be a part of. Lastly, it is assumed that only the top 
5 percent of players in the 
league can be deemed ``superstars'', so they comprise the 95th to 100th 
percentile of the NBA. By this 
method, it is felt that a more realistic legend for PAR has been created as 
opposed to just breaking up the 
data evenly into 5 groups.
